---
title: "Untitled"
output: html_document
---





```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(shiny)
library(choroplethr)
library(choroplethrMaps)
library(usdata)
library(gridExtra)
library(ggthemes)
library(packcircles)
library(treemap)
library(tidyr)
library(usmap)
```


```{r}
url = 'https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv'
df <- read.csv(url)
dim( df )
head( df )
glimpse( df )
```

## Question 1:
As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by the crude mortality of each cause of death.  

```{r}
# preparing the data

# filter for 2010
df_2010 <- df %>%
  filter( Year == 2010 )

dim( df_2010 )
head( df_2010 )
```


```{r}
# create an example plot
df_2010_exp <- df_2010 %>% 
  filter( ICD.Chapter == 'Certain infectious and parasitic diseases' ) 

#bar plot
y_lines = c( 20, 40, 60 )
plt1 <- df_2010_exp %>%
  ggplot( aes( x= reorder(State,Crude.Rate), y= Crude.Rate ) ) +
  geom_col( ) +
  coord_flip() +
  # labs(
  #   title = 'Number of top 5000 companies by State',
  #   x = 'State',
  #   y = '# Companies'
  # ) + 
  theme_classic() +
  geom_hline(yintercept=y_lines, color="white", size=1)
plt1

#choropleth
map_data <- df_2010_exp
map_data$value = map_data[, 6]
map_data$region = tolower( abbr2state( df_2010_exp$State ) )
    
plt2 <- state_choropleth(df = map_data,
                 title = colnames(map_data)[2], 
                 num_colors = 7) + 
  theme(legend.position=c(.9, .23)) +
  #scale_fill_brewer(palette="YlOrBr") +
  labs(title = 'Certain infectious and parasitic diseases',
       subtitle = "2010 US Census",
       caption = "source: https://en.wikipedia.org/wiki/List_of_U.S._states_by_Hispanic_and_Latino_population",
       fill = "Crude Mortality Rate") 

#grid.arrange( plt1, plt2, ncol = 2, widths = c(1,2) )
plt2
```

```{r}
#packcircles
packing <- circleProgressiveLayout(map_data$Crude.Rate, sizetype='area')
map_data <- cbind(map_data, packing)
dat.gg <- circleLayoutVertices(packing, npoints=50)
# Make the plot
ggplot() + 
  # Make the bubbles
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=as.factor(id)), colour = "black", alpha = 0.6) +
  # Add text in the center of each bubble + control its size
  geom_text(data = map_data, aes(x, y, size=Crude.Rate, label = State)) +
  scale_size_continuous(range = c(1,4)) +
  # General theme:
  theme_void() + 
  theme(legend.position="none") +
  coord_equal()
```
I'm not crazy about the circle plot.  let's try another....

```{r}
treemap(map_data,
            index="State",
            vSize="Crude.Rate",
            type="index"
            )
```



```{r}
ui <- fluidPage(
  headerPanel('Mortality Rate Explorer'),
  theme = shinythemes::shinytheme( 'superhero' ),
  sidebarPanel(
    selectInput('icd', 'International Classification of Disease', unique(df_2010$ICD.Chapter), selected='Certain infectious and parasitic diseases'),
  ),
  mainPanel(
    h4('Compare Mortality Rates from Particular International Classification of Disease across Different States'),
    tabsetPanel(
      tabPanel( 'Map View', plotOutput('plot1'), verbatimTextOutput('stats')),
      tabPanel( 'Ranked View', plotOutput('plot2')),
      tabPanel( 'Treemap View', plotOutput('plot3'))
  ))
)

server <- shinyServer(function(input, output, session) {
  
  selectedData <- reactive({
    dfSlice <- df_2010 %>%
      filter(ICD.Chapter == input$icd) %>%
      mutate( value = Crude.Rate,
              region = tolower( abbr2state( State ) ))
  })
  
  output$plot1 <- renderPlot({
    
    dfSlice <- df_2010 %>%
      filter(ICD.Chapter == input$icd)
    
    state_choropleth(df = selectedData(),
                 num_colors = 7) + 
      theme(legend.position=c(.9, .23)) +
      labs(title = input$icd,
           fill = "Crude Mortality Rate") + 
      theme(plot.title = element_text(hjust = 0.5, size = 18, family = 'serif'))
  })
  
  output$stats <- renderPrint({
    dfSlice <- selectedData() %>%
      filter(ICD.Chapter == input$icd)

    summary(dfSlice$Crude.Rate)
  })
  
  output$plot2 <- renderPlot({
    dfSlice <- selectedData() %>%
      filter(ICD.Chapter == input$icd)
    
    ggplot( selectedData(),
            aes( x= reorder(State,Crude.Rate), y= Crude.Rate ) ) +
      geom_col( ) +
      coord_flip() +
      labs( 
        title = 'Ranked Mortality Rate',
        subtitle = input$icd,
        y = "Crude Mortality Rate",
        x = 'State'
        ) + 
      theme_tufte(base_size = 14)
  },
  width = 500,
  height = 800) 
  
  output$plot3 <- renderPlot({
    dfSlice <- selectedData() %>%
      filter(ICD.Chapter == input$icd)
    
treemap(dfSlice,
            index="State",
            vSize="Crude.Rate",
            type="index"
            ) + 
      theme_tufte(base_size = 14)
  },
  width = 500,
  height = 800) 
  
})

shinyApp(ui = ui, server = server)

```
   


## Question 2
Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at the time. Keep in mind that the national average should be weighted by the national population

```{r}
nataves <- df %>%
  group_by( ICD.Chapter, Year ) %>%
  summarise( popsum = sum( Population ), deathsum = sum( Deaths ), n = n() ) %>%
  mutate( nat_crude_death_rate = deathsum/popsum*100000)
head( nataves )
```
CRUDE DEATH RATE is the total number of deaths to residents in a specified geographic area (country, state, county, etc.) divided by the total population for the same geographic area (for a specified time period, usually a calendar year) and multiplied by 100,000.
```{r}
res <- left_join( df, nataves, by = c( "ICD.Chapter", 'Year')) %>%
  mutate( difference = Crude.Rate - nat_crude_death_rate )

head( res )
```
```{r}
resf <- res %>%
  filter(ICD.Chapter == "External causes of morbidity and mortality" , 
         State == 'NE' ) %>%
  mutate( dif_sign = difference > 0 )
ggplot( resf ) +
  geom_line( aes( x = Year, y = nat_crude_death_rate ) ) +
  geom_line( aes( x = Year, y = Crude.Rate, colour = difference > 0 )) +
  annotate( 'text', x = 2001, y = resf$nat_crude_death_rate[1] + 5, color = 'black', size = 4, label = c('National Average')) +
  scale_colour_manual(values = setNames(c('red','blue'),c(T, F))) +
  ylim( 0, max( resf$Crude.Rate, resf$nat_crude_death_rate, 100 ) ) +
  theme_classic() +
  theme(legend.position='none')
```
```{r}
resf %>%
  select( Year, Crude.Rate, nat_crude_death_rate) %>%
  pivot_longer( !Year, names_to = 'group', values_to = 'rate' ) %>%
  ggplot( aes( x = Year, y= rate, fill = group)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme_tufte()
```


```{r}
res %>%
  filter(ICD.Chapter == "External causes of morbidity and mortality", 
         Year == 2010) %>%
      mutate( value = difference,
              region = tolower( abbr2state( State ) ),
              fips = fips( State ) ) %>%
  select( Year, Crude.Rate, nat_crude_death_rate) %>%
  pivot_longer( !Year, names_to = 'group', values_to = 'rate' )
```


```{r}
ui <- fluidPage(
  headerPanel('Mortality Rate Explorer'),
  h4('Compare State-Level Crude Mortality Rates to the National Rates for a given International Classification of Disease (ICD)'),
  theme = shinythemes::shinytheme( 'superhero' ),
  sidebarPanel(
    tags$head(tags$style("#text1{color: white;
                             font-size: 20px;
                         font-style: bold;
                         }"
                     )
    ),
    tags$head(tags$style("#text1b{color: white;
                             font-size: 18px;
                         }"
                     )
    ),
    tags$head(tags$style("#text2{color: white;
                             font-size: 20px;
                         font-style: bold;
                         }"
    )),
    tags$head(tags$style("#text2b{color: white;
                             font-size: 18px;
                         }"
                     )),    
    selectInput('icd', '(ICD)', unique(res$ICD.Chapter), selected='Certain infectious and parasitic diseases'),
    selectInput('state', 'State', unique(res$State), selected='AL'),
    selectInput('year', 'Year', unique(res$Year), selected='1999')
  ),
  mainPanel( 
    tabsetPanel(
      tabPanel( 'Bar Chart per State', textOutput("text1"), textOutput("text1b") , plotOutput('plot1')),
      tabPanel( 'Map View per Year', textOutput("text2"), textOutput("text2b"), plotOutput('plot2'))
  ))
)

server <- shinyServer(function(input, output, session) {
  output$text1 <- renderText({
    "Chose a ICD and State:"
  })
  
  output$text1b <- renderText({
    "observe how a given state's mortality rate compares to the national average across all years for which data is present"
  })

  output$text2 <- renderText({
    "Chose a ICD and Year:"
  })
  
  output$text2b <- renderText({
    "Observe the state-level deviation of mortality rate from the national average for a given year"
  })
  
  selectedData <- reactive({
    dfSlice <- res %>%
      filter(ICD.Chapter == input$icd) %>%
      mutate( value = difference,
              region = tolower( abbr2state( State ) ),
              fips = fips( State ) )
  })
  
  output$plot2 <- renderPlot({
    orange <- "#BB0404"
    blue <- "#0013D3"
    map_dat <- selectedData() %>% filter( Year == input$year )
    plot_usmap(data = map_dat, values = "difference",  color = orange, labels=FALSE) + 
      scale_fill_gradient2( low = blue, mid = "white", high = orange, 
                             name = "Difference", label = scales::comma ) + 
      theme(legend.position = "right") + 
      theme(panel.background = element_rect(colour = "black")) + 
      theme(legend.position=c(.9, .23)) +
      labs(title = input$icd,
           subtitle = "Mortality Rate Difference: National - State",
           fill = "Crude Mortality Rate") +       
      theme(plot.title = element_text(hjust = 0.5, size = 18, family = 'serif'),
            plot.subtitle = element_text(hjust = 0.5, size = 16, family = 'serif')) #+
      #theme_classic()
  })
  
  
  output$plot1 <- renderPlot({
    dfSlice <- selectedData() %>%
      filter( State == input$state ) %>%
      select( Year, Crude.Rate, nat_crude_death_rate) %>%
      pivot_longer( !Year, names_to = 'group', values_to = 'rate' )
    
    ggplot(dfSlice, aes( x = Year, y= rate, fill = group)) +
      geom_bar(stat="identity", position=position_dodge()) +
      labs(
        title = paste0( input$state, ': ', input$icd  ),
        subtitle = "State-level vs National Crude Mortality",
        y = "Mortality Rate",
        x = 'Year'
      ) +
      theme_classic() +
      scale_fill_discrete(labels = c( input$state,'National Average' )) +
      theme(plot.title = element_text(hjust = 0.5, size = 18, family = 'serif'),
            plot.subtitle = element_text(hjust = 0.5, size = 16, family = 'serif'))  +
      guides(fill=guide_legend(title=""))
  })
  
  
})

shinyApp(ui = ui, server = server)

```